Intersection Types, Sub-formula Property, and the Functional Character of the Lambda Calculus

### 2018-08-08 DenotSem/determ.tex
check

\newcommand{\lam}[1]{\lambda #1.\,}
\newcommand{\app}[0]{\;}
\newcommand{\IF}[0]{\textbf{if}\;}
\newcommand{\THEN}[0]{\;\textbf{then}\;}
\newcommand{\ELSE}[0]{\;\textbf{else}\;}
\newcommand{\ESEM}[1]{\mathcal{E}[\![ #1 ]\!]}
\newcommand{\FSEM}[1]{\mathcal{F}[\![ #1 ]\!]}
\newcommand{\SET}[1]{\mathcal{P}(#1)}

@@

@@

@@

>>>>

\newcommand{\lam}[1]{\lambda #1.\,}
\newcommand{\app}[0]{\;}
\newcommand{\IF}[0]{\mathrm{if}\;}
\newcommand{\THEN}[0]{\;\mathrm{then}\;}
\newcommand{\ELSE}[0]{\;\mathrm{else}\;}
\newcommand{\ESEM}[1]{\mathcal{E}[\![ #1 ]\!]}
\newcommand{\FSEM}[1]{\mathcal{F}[\![ #1 ]\!]}
\newcommand{\SET}[1]{\mathcal{P}(#1)}
\newcommand{\WF}[1]{\mathsf{wf}(#1)}
\newcommand{\UP}[1]{\mathord{\uparrow} #1}

@@

\section{Intersection Types to the Rescue}

@@

\section{Subtyping and the Sub-formula Property}

@@

\section{The Home Stretch}

Armed with a good definition of intersection types, I turned back to
the proof of the Join Theorem, but first I needed to adjust my
denotational semantics to use intersection types. For this we'll need
the definition of well formed types that we alluded to earlier.

\begin{gather*}
  \frac{}{\WF{n}}
  \qquad
  \frac{\WF{A} \quad \WF{B}}{\WF{A \to B}}
  \qquad
  \frac{A \sim B \quad \WF{A} \quad \WF{B}}{\WF{A \sqcap B}}
\end{gather*}

We also introduce the following notation for all the well-formed types
that are super-types of a given type.
\[
   \UP{A} \stackrel{def}{=} \{ B\mid A <: B \text{ and } \WF{B} \}
\]

We shall represent variables with de Bruijn indices, so an environment
$\rho$ is a sequence of types.  The denotational semantics of the CBV
lambda calculus is defined as follows.
\begin{align*}
  \ESEM{n}\rho &= \UP{n} \\
  \ESEM{x}\rho &=  \IF x < |\rho| \THEN \UP{\rho_k} \ELSE \emptyset \\
  \ESEM{\lam{x}e }\rho &= \{ A \mid \WF{A} \text{ and } \FSEM{A}(e,\rho) \} \\
  \ESEM{e_1\app e_2}\rho &= \{ C \mid
    A \in \ESEM{e_1}\rho,
    B \in \ESEM{e_2}\rho,
    A <: B \to C, \text{ and } \WF{C} \} \\
  \ESEM{f(e_1,e_2)}\rho &=
    \{ \}
\end{align*}
